\begin{abstract}
This thesis employs X-ray scattering to study the structure of two different stacked
lipid membrane systems.
The first part reports the effect on lipid bilayers of the Tat peptide
Y$_{47}$GRKKRRQRRR$_{57}$ from the HIV-1 virus transactivator of translation (Tat) protein.
Synergistic use of low angle X-ray scattering (LAXS) and atomistic molecular dynamics (MD)
simulations indicated Tat peptide binding to neutral dioleoylphosphatidylcholine (DOPC)
headgroups. This binding induced the nearby lipid phosphate groups to move 3 \AA\
closer to the bilayer center. Many of the positively charged guanidinium components of
the arginines were as close to the bilayer center as the locally thinned lipid phosphate 
groups. Analysis of LAXS from DOPC, DOPC/dioleoylphosphatidylethanolamine (DOPE),
DOPC/dioleoylphosphatidylserine (DOPS), and a mimic of the nuclear membrane
indicated that the Tat peptide decreased the bilayer bending modulus $K_c$
and increased the area per lipid, facilitating Tat membrane translocation.
Although a mechanism for translocation remains elusive, this study suggests that
Tat translocates from the headgroup region.

The second study presents the structure of the asymmetric ripple phase formed
by dimyristoylphosphatidylcholine. 
We determined the most detailed ripple phase structure by combining synchrotron 
LAXS and wide angle X-ray scattering (WAXS) from highly aligned multilamellar samples.
High intensity LAXS revealed 52 reflections. We derived three corrections 
to calculate the X-ray form factors from the measured intensities.
Analysis of high quality LAXS provided a high resolution two-dimensional 
electron density map, demonstraing that the ripple major arm was consistent 
with the gel $L_{\beta'}$ phase and showing clear distinction between the major 
and minor arm structure, supporting the coexistence of different molecular organizations
in the ripple phase.
The minor arm electron density profile was qualitatively consistent
with interdigitated chain packing previously proposed by MD simulations.
Analysis of high resolution near grazing incidence wide angle X-ray scattering (nGIWAXS)
with the thin rod model developed for the ripple major arm showed that
hydrocarbon chains were tilted parallel to the ripple 
plane by 18.4\textdegree\ with respect to the bilayer local normal, 
toward the next nearest neighbor similarly to the gel $L_{\beta F}$ phase.
By measuring the Bragg rod lengths in transmission WAXS, we determined that 
chains in the opposing leaflets were coupled coherently scattering units.
The LAXS and WAXS results together indicated that chains in the major arm
were shorter by 1.3 \AA\ compared to the $L_{\beta'}$ phase, suggesting a
gouche-trans-gouche kink in the ripple major arm.
In contrast to the LAXS analysis, the measured nGIWAXS was consistent with 
disordered chains similarly to the fluid $L_\alpha$ phase.

\end{abstract}
